---
title: "Explore DwC-DP"
date: today
format: 
  html: 
    df-print: paged
    toc: true
editor: source
---

## Explore the DwC-DP tables

You can load all the RDA files like below and play around with the tables as demonstrated in this page. Any pull request is also welcomed.

```{r load rda}
#| warning: false
#| results: hide

library(here)
library(tidyverse)

# Load all the RDA files
rda_files <- list.files(path = here("data", "output", "rda"), pattern = "\\.rda$", full.names = TRUE)
lapply(rda_files, load, envir = .GlobalEnv)
```

## Community measurements 

Does DwC-DP handles community measurements better than DwCA?

### DwCA

In DwCA, we will have to have all these as Occurrence records so that measurements like "Standard Length" and "Total Length" of each individual in emof has an Occurrence record to point to. However, this does not warrant user to add the organismQuantity of all individuals together.

occurrenceID | eventID | scientificName | lifeStage | organismQuantity | organismQuantityType
---|---|---|---|---|---
BROKE_WEST_RMT_051_RMT8_217697 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 28 | individuals
AAV3FF_00231 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00232 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00233 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00234 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00235 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00236 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00237 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00238 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00239 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00240 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual
AAV3FF_00241 | BROKE_WEST_RMT_051_RMT8 | Electrona antarctica | larvae | 1 | individual

: Occurrence table in DwCA {.responsive .striped .bordered}

### DwC-DP

The Occurrence and Material are separate concepts now. We do not need to have a separate Occurrence just to indicate the total catch of a taxon. We can have a single Occurrence (total catch) with multiple Materials (individual) to indicate the total catch and preserved individuals. The measurements of the individuals (Assertions) can be directly linked to the Materials via the Material Assertion table.

```{r community measurement}
event_id <- "BROKE_WEST_RMT_051_RMT8"
taxon_id <- "urn:lsid:marinespecies.org:taxname:217697"

# total catch of electrona antarctica from BROKE_WEST_RMT_051_RMT8
occurrence %>% filter(eventID == event_id & taxonID == taxon_id) %>% 
  select(eventID, occurrenceID, scientificName, lifeStage, organismQuantity, organismQuantityType)

# all preserved electrona antarctica from BROKE_WEST_RMT_051_RMT8
mat <- material %>% filter(eventID == event_id & taxonID == taxon_id) %>% 
  select(materialEntityID, eventID, materialEntityType, scientificName, preparations)
mat

# all measurements of individual electrona antarctica from BROKE_WEST_RMT_051_RMT8
mat %>% left_join(material_assertion, by = "materialEntityID") %>% 
  select(materialEntityID, assertionType, assertionValueNumeric, assertionValue, assertionUnit)
```


## Body part measurements

Very often, we received dataset with measurements performed on a specific body part of an organism. Example: <https://www.gbif.org/occurrence/3344249657> A specific body part/sample of an organism is not an Occurrence nor an Event, so it is difficult to model with DwCA.

### DwCA

Currently, I modeled it using eMoF pointing to the Occurrence with body part in `measurementRemarks`. Specifying the body part can be embedded in a matrix of a NERC vocabulary but it is not practical to mint NERC for every body part and body part is specific to a taxon. We also want to consider the different 

| occurrenceID              | scientificName                          |
|---------------------------|-----------------------------------------|
| SO_Isotope_1985_2017_1013 | Glabraster antarctica (E.A.Smith, 1876) |

: Occurrence table in DwCA {.responsive .striped .bordered}


| occurrenceID | measurementType | measurementValue | measurementUnit | measurementRemarks |
|---------------|---------------|---------------|---------------|---------------|
| SO_Isotope_1985_2017_1013 | The carbon elemental content measured in the tegument of the considered sea star specimen, expressed in relative percentage of dry mass | 12.28 | relative percentage of dry mass | tegument |

: eMoF table in DwCA {.responsive .striped .bordered}

### DwC-DP

Relationship between Materials can be specified through the `derivedFromMaterialEntityID` field. Example of a krill eaten by a fish can be modeled within a single `Material` table.

```{r material relationship}
material %>% filter(str_starts(materialEntityID, "AAV3FF_00025")) %>% 
  select(materialEntityID, derivedFromMaterialEntityID, materialEntityType, scientificName, materialEntityRemarks)
```

## Non-detections 

### DwCA

Before [Humboldt Extension](https://eco.tdwg.org/) was developed, non-detections are represented as an [Occurrence record with occurrenceStatus = absent](https://www.gbif.org/occurrence/search?occurrence_status=absent) in DwCA.

When Humboldt Extension comes along, [non-detections can be inferred by looking at the target scopes that do not have an Occurrence record](https://raw.githack.com/biodiversity-aq/humboldt-for-eco-survey-data/main/src/mapping.html) if the target scope was fully reported.

### DwC-DP

In DwC-DP, non-detections can be represented as an Occurrence with occurrenceStatus = absent, just like in DwCA. Similarly, non-detections can also be inferred by looking at the SurveyTarget that do not have an Occurrence record.

```{r non-detections}

```


## Trade off between expressiveness of DwC-DP and simplicity of DwCA?
