---
title: "Explore DwCA v2"
date: today
format: 
  html: 
    df-print: paged
editor: visual
---

## Questions

### How can I explore the tables for DwCA v2 from this mapping exercise?

You can load the tables from the RDA files like below and play around with the tables as demonstrated in this page. Any pull request is also welcomed.

```{r load rda}
#| warning: false
#| message: false
library(here)
library(tidyverse)

load(here("data", "output", "rda", "agent.rda"))
load(here("data", "output", "rda", "event.rda"))
load(here("data", "output", "rda", "identification.rda"))
load(here("data", "output", "rda", "material.rda"))
load(here("data", "output", "rda", "material_gathering.rda"))
load(here("data", "output", "rda", "survey_target.rda"))
load(here("data", "output", "rda", "survey_target_abundance.rda"))
```

### Does DwCA v2 solve community measurements issue?

Yes! Example: Total count of electrona antarctica is recorded in `SurveyTargetAbundance` table, only a subset was preserved and recorded in this dataset. The preserved fish is recorded in the `Material` table.

```{r}
survey_target_abundance %>% filter(surveyID == "BROKE_WEST_RMT_051_8_S" & observedTaxon == "Electrona antarctica")

material %>% filter(materialGatheringID == "BROKE_WEST_RMT_051_8_MG" & scientificName == "Electrona antarctica")
```

### Can relationship between Materials be expressed in DwCA v2?

Very often, we received dataset with measurements performed on a specific body part of an organism. Example: <https://www.gbif.org/occurrence/3344249657>

#### DwCA

Currently, I modeled it using eMoF pointing to the Occurrence with body part in measurementRemarks. Specifying the body part can be done through NERC vocabulary. However, it is a hassle to mint NERC for every body part and body part is specific to a taxon.

**Occurrence table**

| occurrenceID              | scientificName                          |
|---------------------------|-----------------------------------------|
| SO_Isotope_1985_2017_1013 | Glabraster antarctica (E.A.Smith, 1876) |

**eMoF table**

| occurrenceID | measurementType | measurementValue | measurementUnit |
|------------------|------------------|------------------|------------------|
| SO_Isotope_1985_2017_1013 | The carbon elemental content measured in the tegument of the considered sea star specimen, expressed in relative percentage of dry mass | relative percentage of dry mass | tegument |

#### DwCA v2

Yes! Relationship between Materials can be specified through the `Material` table. Example of a krill eaten by a fish can be modeled within a single `Material` table

```{r}
material %>% filter(str_starts(materialEntityID, "AAV3FF_00025"))
```

### How non-detections are represented in DwCA v2?

### Trade off between expressiveness of DwCA v2 and simplicity DwCA?
